\documentclass[10pt,twocolumn,letterpaper]{article}
\usepackage[margin=0.85in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,fit,backgrounds,calc}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pifont}

% Tighter spacing
\setlength{\columnsep}{0.25in}
\setlength{\parskip}{2pt}

% ── Placeholder macros for headline result numbers ─────────────────────────
% These are overwritten by figs/inject_results.py once predictions are scored.
\newcommand{\PLBASEDISH}{39.4\%}
\newcommand{\PLDISTDISH}{86.4\%}
\newcommand{\PLBASECEX}{46.3\%}
\newcommand{\PLDISTCEX}{90.3\%}
\newcommand{\PLBASEMM}{39.4\%}
\newcommand{\PLDISTMM}{86.4\%}
\newcommand{\PLBASESETIOU}{47.0\%}
\newcommand{\PLDISTSETIOU}{92.1\%}
\newcommand{\PLBASEDELTAN}{$-0.54$}
\newcommand{\PLDISTDELTAN}{$-0.04$}
\newcommand{\PLBASEABSDELTAN}{0.57}
\newcommand{\PLDISTABSDELTAN}{0.10}
\newcommand{\PLBASEZERODET}{37.8\%}
\newcommand{\PLDISTZERODET}{0.0\%}
\newcommand{\PLDISHDELTA}{+47.0~pp}
\newcommand{\PLNEVAL}{596}

\title{\vspace{-2em}\textbf{TahamajaNet:\ A 72B$\rightarrow$7B Distillation Pipeline\\with Cardinality-Aware Fine-Tuning and an End-to-End Semantic Metric for Cafeteria Tray Analysis}}

\author{Anonymous Authors\\KFUPM Restaurant Project}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\noindent
We present \textbf{TahamajaNet}, a three-stage detection pipeline that turns
a single cafeteria-tray image into the multiset of food items needed to
generate a bill, and we identify the central failure mode of generative-VLM
detection on this task: \emph{for VLM-as-detector, syntax and localization
are not the bottleneck; cardinality is}. We make four contributions.
\textbf{(i)~A 72B$\rightarrow$7B distillation pipeline}: a 72B teacher
labels $3{,}702$ trays once offline (Qwen2.5-VL-72B-AWQ\,$\rightarrow$\,SAM3\,$\rightarrow$\,DINOv2/SigLIP cluster
review) and a 7B student runs the same three stages online. \textbf{(ii)~A
cardinality-aware structural loss} that addresses an asymmetric overcount
regression we measure quantitatively: matched bbox IoU climbs from $0.32$ to
$0.90$ across training while exact-count accuracy drops from $86.8\%$ to
$53.5\%$. \textbf{(iii)~Dish-Correct}, an end-to-end semantic metric
(count-exact $\wedge$ class-multiset-match) that aligns with bill correctness
and forgives bbox geometry. \textbf{(iv)~The v3.2 dataset}, a $4{,}629$-item,
$32$-class, hand-curated Saudi cafeteria-tray dataset, released with splits.
On $596$ held-out images, the distilled 7B reaches
\textbf{\PLDISTDISH{}} dish-correct, a \textbf{\PLDISHDELTA{}} absolute lift
over the off-the-shelf $7$B baseline (zero-detection rate
$\PLBASEZERODET\!\rightarrow\!\PLDISTZERODET$, per-item recall
$45.4\%\rightarrow90.6\%$).
\end{abstract}

%-----------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

Generating an itemized bill from a single photo of a cafeteria tray is a
deceptively hard vision problem. The output is not a set of pixel-accurate
masks, nor a ranked list of detection scores: it is a \emph{multiset of
food classes} priced by a downstream lookup. A single missed item or a single
duplicated detection is a wrong bill. A bbox that overlaps the ground truth
at IoU 0.5 versus 0.9 is, at the level of revenue, indistinguishable.

This semantic objective interacts poorly with two off-the-shelf options.
Large vision-language models (VLMs) at the 70B parameter scale will reliably
\emph{describe} a tray image but are too slow and too expensive for inference
on every dish (target latency: $<2{\,}000$ ms per image). Small open-set
detectors are fast but lack the dish-level concept ontology required to merge
toppings with their base dish, separate co-located items, and ignore plates
and utensils.

We bridge these regimes with four contributions.

\textbf{(1)~A 72B$\rightarrow$7B distillation pipeline (\S\ref{sec:dataengine},
\S\ref{sec:student}).} A Qwen2.5-VL-72B-AWQ teacher labels 3{,}702 cafeteria
images via a structured-JSON prompt, followed by SAM3 segmentation, DINOv2 +
SigLIP clustering, and human cluster curation. The resulting 4{,}938-item
dataset (32 classes) trains a Qwen2.5-VL-7B student via LoRA on the language
model and an unfrozen vision encoder. The teacher pipeline runs once,
offline; the student runs at every inference.

\textbf{(2)~Cardinality-aware fine-tuning (\S\ref{sec:cardinality}).} We
observe a clear \emph{cardinality-localization tradeoff}: standard supervised
fine-tuning improves bbox geometry while degrading exact instance count, with
overcount the dominant error mode. We introduce a structural count loss
that penalises $|\hat{n} - n|$ during training and arrests the regression.
Our central claim: \emph{for generative VLM-based detection, syntax and
localization are not the bottleneck; cardinality is}.

\textbf{(3)~Dish-Correct (\S\ref{sec:metric}).} We define an end-to-end
semantic evaluation metric whose hierarchy --- count-exact, class set-IoU,
class multiset match, dish-correct --- aligns with the bill-generation
objective and forgives bbox geometry. We argue per-stage IoU is a misleading
proxy in this regime, and we report all results on Dish-Correct.

\textbf{(4)~The v3.2 dataset, public release.} A 4{,}629-item, 32-class
hand-curated Saudi cafeteria-tray dataset --- the first of its scale and
domain that we are aware of. The release reflects significant compute
(96-minute Qwen-VL-72B vLLM run on 80\,GB GPUs, plus DINOv2-large +
SigLIP embedding for $\sim$5{,}000 crops, plus HDBSCAN over the joint
embedding) and significant human time: a curator inspected and labeled
\emph{every cluster centroid}, collapsing 467 raw HDBSCAN clusters into
the 32 production classes, and recovered 1{,}344 noise-bin items via
visual review. Train / dev / test splits and the canonical 80/10/10 seed
are part of the release.

%-----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

\paragraph{VLM-based detection.} Recent instruction-tuned vision-language
models (Qwen-VL, Qwen2.5-VL, LLaVA, InternVL, Florence-2, PaLI-X) cast
bounding-box generation as a JSON-output token-prediction task: the model
emits a structured response listing detected regions with coordinates and
optional descriptors. This paradigm is attractive because the same prompt
template generalises across object categories without needing a fixed-class
detection head, and because the resulting boxes can be coupled with
language-grounded captions in a single forward pass. The trade-off is
twofold: (i)~there is no explicit objectness prior over the predicted
\emph{number} of items, and (ii)~bbox coordinates are emitted as text
tokens that compete for attention with the region-naming tokens. We
inherit both of these trade-offs and address (i) directly with a structural
count loss (\S\ref{sec:cardinality}); (ii) we sidestep by training under a
\emph{bbox-only} schema that strips region names from the supervision
target, leaving fine-grained classification to a downstream embedding head.

\paragraph{Knowledge distillation.} The bulk of distillation literature ---
Hinton's response-matching, FitNets feature-matching, attention transfer,
self-distillation --- assumes a single \emph{model} teacher whose internal
states or output distributions are aligned with those of a student. Our
setup is structurally different: the 72B teacher is a \emph{labeling
pipeline}, and the supervision target is the teacher's deterministic
post-processed output (class-collapsed bboxes after SAM3 mask refinement
and human cluster review) rather than soft logits. There is no
hidden-state match, no temperature-softmax KL term, no auxiliary loss
on intermediate features. This makes the teacher swap-replaceable: any
labeling pipeline that produces structured-JSON detections at acceptable
quality could substitute for Qwen-VL-72B-AWQ. The price we pay is that we
cannot exploit the teacher's uncertainty signal --- a soft logit might
have told us \emph{this is 60\% rice, 40\% biryani}, and our cluster-collapsed
hard label cannot. We treat the resulting label noise as in-distribution.

\paragraph{Multi-stage detection pipelines.} Two-stage R-CNN-family
detectors (Faster R-CNN, Mask R-CNN) decouple region proposal from
classification. End-to-end DETR and its descendants (Deformable DETR,
DINO) use a single transformer over learned object queries. SAM and
SAM2/SAM3 separated mask refinement from prompt generation, enabling
compositional pipelines like Grounded-SAM (open-vocabulary box-from-text
$\rightarrow$ SAM mask). Our pipeline is closest in spirit to
Grounded-SAM but inverts the roles: the box generator is fine-tuned for a
closed taxonomy, the segmenter is frozen, and the classification step is
explicitly retrieval (cosine over a fixed reference bank) rather than a
learned head. The key novelty is the \emph{symmetric} reuse of the same
three-stage decomposition at both train-time labeling (72B) and inference
(7B), at $\sim$10$\times$ different scales --- which makes the train/test
distribution mismatch easier to reason about and lets us re-use
infrastructure (the SAM3 segmenter, the DINOv2 vector DB) between the two
phases verbatim.

\paragraph{Open-vocabulary and food-specific detection.} Food recognition
benchmarks (Food-101, Food-2K, ETHZ Food-101) are predominantly
classification, not detection. The few detection-grade food datasets
(UEC-Food100, UNIMIB, ETH Mensa) are small (under $5{,}000$ images each)
and skewed toward European/Asian cuisines. None of them, to our knowledge,
contain the class taxonomy or cuisine specificity required for Saudi
cafeteria-tray analysis: kabsa rice, masakaah, asidah/Saudi porridge,
chicken/meat koftas, hummus (cling-wrapped), and the wrapped/sealed
dairy sub-categories absent from international datasets. Our v3.2 release
fills this gap.

\paragraph{Cardinality-aware evaluation.} Standard detection metrics ---
COCO mAP, mean IoU, F1@0.5 --- reward bbox geometry and ranking quality
but are blind to absolute item count. Recent work in dense crowd counting
(MCNN, CSRNet) and instance-level industrial inspection has argued for
cardinality-explicit objectives. Dish-Correct generalises this argument
to the multi-class, multi-instance, item-level-billing regime: at the
output of the pipeline, the only thing that matters is whether the
predicted multiset of class labels matches the ground-truth multiset. Bbox
geometry is just an internal coordinate, not a deliverable.

%-----------------------------------------------------------------------------
\section{Data Engine: The Teacher Pipeline}
\label{sec:dataengine}

\input{sections/03_data_engine.tex}

%-----------------------------------------------------------------------------
\section{Student Pipeline: TahamajaNet Inference}
\label{sec:student}

\input{sections/04_student.tex}

%-----------------------------------------------------------------------------
\section{Cardinality-Aware Fine-Tuning}
\label{sec:cardinality}

\input{sections/05_cardinality.tex}

%-----------------------------------------------------------------------------
\section{Exact Subset Accuracy (Dish-Correct): An End-to-End Semantic Metric}
\label{sec:metric}

\input{sections/06_metric.tex}

%-----------------------------------------------------------------------------
\section{Experiments}
\label{sec:experiments}

\input{sections/07_experiments.tex}

%-----------------------------------------------------------------------------
\section{Discussion and Future Work}
\label{sec:discussion}

\input{sections/08_discussion.tex}

%-----------------------------------------------------------------------------
\section*{Acknowledgments}
KFUPM Restaurant Project. Compute via vast.ai (H100 SXM 80GB).

\end{document}
